---
title: "From Python to R and back....!"
---

# Overview

Before I start, I wanted to test the combination of Python and R in the same `.qmd` file.

Apparently, it's very simple — just add `library(reticulate)` in an R cell and you're all set.

To pass objects (e.g., variables, data frames, and so on...) between the two languages, you need to reference the *variable* as `r.[variable]` in Python or `py$[variable]` in R.

So where do the problems start? When you try to insert Python inline code inside markdown text.\
In short, even though your `.qmd` document contains both R and Python executable code cells, it can execute **only** R inline code.\\ To bypass this, there are a bunch of `reticulate::` functions that can help. The two main ones are:

-   `reticulate::py_str()` — roughly equivalent to `print(str(...))`

-   `reticulate::py_eval()` — executes a Python expression that returns numbers, strings, lists, or dicts (can be problematic with Python objects like `type`, classes, or modules)

**Example:** ``{r} reticulate::py_eval("str(type(x))")`` — returns the type of variable `x` as a string

# Examples

It's good practice to start by loading the `reticulate` package:

```{r}
# <R>

# Load reticulate package
library(reticulate)
```

## Passing variable from Python to R

```{python}
# <Py>

# Create Python variable py_x
py_x = 10
print(py_x)
```

Type of py_x: *`{r} reticulate::py_eval("str(type(py_x))")`*

```{r}
# <R>

# Passing Python variable py_x to R 
r_x <- py$py_x
print(r_x)
```

Class of r_x: *`{r} print(class(r_x))`*

## Passing variable from R to Python

```{r}
# <R>

# Create R variable r_y 
r_y <- 20
print(r_y)
```

Class of r_y: *`{r} print(class(r_y))`*

```{python}
# <Py>

# Passing the R variable r_y to Python 
py_y = r.r_y
print(py_y)
```

Type of py_y: `{r} reticulate::py_eval("str(type(py_y))")`

## Passing DataFrame from Python to R

```{python}
# <Py>

# Create a pandas df from .csv file
import pandas as pd

url = "https://raw.githubusercontent.com/mwaskom/seaborn-data/refs/heads/master/flights.csv"

py_flights = pd.read_csv(url)
py_flights.head()
```

Type of py_flights: *`{r} reticulate::py_eval("str(type(py_flights))")`*

```{r}
# <R>

# Passing the pandas df to R data.frame
r_flights <- py$py_flights
head(r_flights)
```

Class of r_flights: `{r} print(class(r_flights))`